---
title: "create_cemdap_input"
output: html_document
date: '2022-04-29'
---

## import library
```{r}
library(dplyr)
```

## read cemdap test data
```{r}
setwd("G:/CE-CERT/CEMDAP/TestData/")
h = read.table("households.dat")
p = read.table('persons.dat')
```

## read riverside demorgraphic data
```{r setup, include=FALSE}
setwd("G:/CE-CERT/CEMDAP/Riv_data")
house = read.csv("households.csv")
person = read.csv("people.csv")
#person = person[order(person$hh_id),]
#hhid = unique(person$hh_id)
#house =  as.data.frame(cbind(house,hhid))
#write.csv(house,'households.csv',row.names = F)
#write.csv(person,'people.csv', row.names = F)

blocks = read.csv("riv_blocks.csv")
bg = ifelse(nchar(blocks$TRACTCE10) == 5, paste0("6065","0",blocks$TRACTCE10,blocks$BLKGRPCE10),
                  paste0("6065",blocks$TRACTCE10,blocks$BLKGRPCE10))
bg = unique(bg)

house$bg = ifelse(nchar(house$tract) == 5, paste0(house$state,"0",house$county,"0",house$tract,house$block.group),
                  paste0(house$state,"0",house$county,house$tract,house$block.group))

house = house[which(house$bg %in% bg),]
person = person[which(person$hh_id %in% house$hhid),]
```

## household and person selected
```{r}
n = 111383 ## choose n households
house1 = house[1:n,]
m=sum(house1$NP)
person1 = person[1:m,]
```

# read zone id
```{r}
zone_refer = read.csv("zone_refer.csv")
zone_refer$bg=as.character(zone_refer$bg)
house1 = left_join(house1,zone_refer,by="bg")
```


# for household # part I
```{r}
hhid = house1$hhid
nadult = person1 %>% group_by(hh_id) %>% summarise(adult=sum(AGEP>=18)) 
nadult = nadult$adult
nveh = house1$VEH
#hometsz = sample(seq(2000,2200,1),n,replace = T)
nchild = person1 %>% group_by(hh_id) %>% summarise(child=sum(AGEP<18))
nchild = nchild$child
```

# for household # part II
```{r}
hhstr_direct=c("40","10","11","20","22","24","30","50")
#hhstr_direct=c("31")
hhstr_indirect=c("41","42","43","51","53","21","23","31","32","33")
hhstr = c(hhstr_direct,hhstr_indirect)
npattern = paste0(nadult,nchild)
#plot(as.data.frame(table(npattern)))
#write.csv(as.data.frame(table(npattern)),"npattern.csv",row.names = F)
test = house[which(npattern %in% hhstr),]
sum(test$NP)
```

# for hhstruct = 31 only
```{r}
change21 = which(npattern=="31")
change21_id=house1[change21,]$hhid

person11=person1[which(person1$hh_id %in% change21_id),]
person1 = person1[-which(person1$hh_id %in% change21_id),]

person11 = person11[-which(person11$AGEP<18),]

person11$RELP = ifelse(person11$RELP %in% c(2,6),10,person$RELP)
```

# for INDIRECT hhstruct 
```{r}
change = which(npattern %in% hhstr_indirect)
change_id=house1[change,]$hhid

person12=person1[which(person1$hh_id %in% change_id),]
person1 = person1[-which(person1$hh_id %in% change_id),]


person12$RELP = ifelse(person12$RELP %in% c(2,6),10,person$RELP)

person12$AGEP = ifelse(person12$AGEP<18,18,person12$AGEP)
```

#------------above is the process to 1. test useful hhstruct. 2. split direct households and indirect households
# for households part III, consider direct households
```{r}
house_direct = house1[which(npattern %in% hhstr),]
house_direct_id = house_direct$hhid
person1 = as.data.frame(rbind(person1,person11,person12))
person1 = person1[order(person1$hh_id),]
person_direct = person1[which(person1$hh_id %in% house_direct_id),]
```

# hh structure generation
```{r}
hhstr_direct=c("10","11","20","21","23","24","22","30","31","32","50","60","70","80","40")
hhstruct_standard = as.data.frame(matrix(c(1,2,9,10,12,13,15,16,17,18,16,16,16,16,16,hhstr_direct),byrow = F,ncol = 2))
names(hhstruct_standard)=c("hhstr","pattern")
```

# PART I
```{r}
hhid = house_direct$hhid
nadult = person_direct %>% group_by(hh_id) %>% summarise(adult=sum(AGEP>=18)) 
nadult = nadult$adult
nveh = house_direct$VEH
#hometsz = sample(h$V4,length(hhid),replace = T)
hometsz = house_direct$zone_index
nchild = person_direct %>% group_by(hh_id) %>% summarise(child=sum(AGEP<18))
nchild = nchild$child
npers = nadult + nchild
temp1 = as.data.frame(cbind(nadult,nchild))
temp1$pattern = paste0(nadult,nchild)
temp1 = left_join(temp1,hhstruct_standard, by="pattern")
struct = temp1$hhstr
number_employed_adult = person_direct %>% group_by(hh_id) %>% summarise(nemadult = sum(AGEP>=18 & ESR %in% c(1,2,4,5)))
number_employed_adult=number_employed_adult$nemadult
naunemp = nadult - number_employed_adult
#table(naunemp)
multadult = ifelse(nadult>=2,1,0)
vehbylic = ifelse(nadult==0,0,nveh/nadult)
sperson = ifelse(struct==1,1,0)
sparent = rep(1,dim(house_direct)[1])
couple = ifelse(struct==9,1,0)
nuclear = ifelse(struct %in% c(2,10,12,13,17,18),1,0)
ohhtype = ifelse(struct %in% c(15,16),1,0)
numlic = nadult
zeroch = ifelse(nchild==0,1,0)
onech = ifelse(nchild==1,1,0)
twoch = ifelse(nchild>1,1,0)
naemp = number_employed_adult
zeroemp = ifelse(naemp == 0,1,0)
oneemp = ifelse(naemp == 1,1,0)
twoemp = ifelse(naemp>1,1,0)
nastu = nadult-naunemp-naemp
ncnotstu = person_direct %>% group_by(hh_id) %>% summarise(nc=sum(AGEP<18 & SCH %in% c(NA,1)))
ncnotstu=ncnotstu$nc
ncstu = nchild - ncnotstu
ownhome = ifelse(naemp == 1,1,1)
sfdunit = rep(1,dim(house_direct)[1])
sfaunit = rep(1,dim(house_direct)[1])
aptunit = rep(1,dim(house_direct)[1])
othunit = rep(1,dim(house_direct)[1])
hhincome = house_direct$HINCP
```


## combine variables into households table
```{r}
setwd("G:/CE-CERT/CEMDAP/test1/")
households_generated =  as.data.frame(cbind(hhid,nadult,nveh,hometsz,nchild,struct,naunemp,hhincome,zeroch,npers,
                                            multadult,vehbylic,sperson,sparent,couple,nuclear,ohhtype,numlic,onech,
                                            twoch,naemp,zeroemp,oneemp,twoemp,nastu,ncnotstu,ncstu,ownhome,sfdunit,
                                            sfaunit,aptunit,othunit))
households_generated[is.na(households_generated)] = 0
households_generated=lapply(households_generated,as.numeric)
dim(households_generated)
write.table(households_generated,"households.dat",sep = "\t",col.names = F,row.names = F)
### the warning is OK
```

```{r}

```

## direct persons
```{r}
hhid = person_direct$hh_id
perid = unlist(lapply(npers, seq))
aemp = ifelse(person_direct$AGEP>=18 & person_direct$ESR %in% c(1,2,4,5),1,0)
stu = ifelse(person_direct$AGEP<18 & person_direct$SCH %in% c(2,3),1,0)
Lic = ifelse(person_direct$AGEP>=18,1,0)
#worktsz = sample(hometsz,length(hhid),replace = T)
#worktsz = sample(house_direct$zone_index,length(hhid),replace = T)
#worktsz = rep(2200,length(hhid))
worktsz = ifelse(aemp==1,sample(job_by_zone$zone_index,length(hhid),prob = job_by_zone$job,replace = T),-99)

#schooltsz = sample(hometsz,length(hhid),replace = T)
#schooltsz = sample(house_direct$zone_index,length(hhid),replace = T)
schooltsz = rep(hometsz,npers)
schooltsz = ifelse(stu==1,schooltsz,-99)
female = ifelse(person_direct$person_sex == 'male',0,1)
age = person_direct$AGEP
parent = ifelse(person_direct$RELP == 6,1,0)
cauc =ifelse(person_direct$race == "white",1,1)
afamer = ifelse(person_direct$race == "white",0,0)
asian = ifelse(person_direct$race == "white",0,0)
hisp = ifelse(person_direct$race == "white",0,0)
othrace = ifelse(person_direct$race == "white",0,0)
male = ifelse(person_direct$person_sex == 'male',1,0)

hhid_temp = house_direct$hhid
temp = as.data.frame(cbind(hhid_temp,vehbylic))
hhid = as.data.frame(hhid)
temp1 = left_join(hhid,temp,by=c('hhid'='hhid_temp'))

pvehavbl = ifelse(stu == 1, 0, temp1$vehbylic)

presch = ifelse(person_direct$AGEP<18 & person_direct$SCHL == 2,1,0)
kgtog4 = ifelse(person_direct$AGEP<18 & person_direct$SCHL %in% c(3,4,5,6,7),1,0)
g5tog8 = ifelse(person_direct$AGEP<18 & person_direct$SCHL %in% c(8,9,10,11),1,0)
g9orhigh = ifelse(person_direct$AGEP<18 & person_direct$SCHL >= 12,1,0)
g8orlow = ifelse(person_direct$AGEP>=18 & person_direct$SCHL <= 11,1,0)
g9tog12 = ifelse(person_direct$AGEP>=18 & person_direct$SCHL %in% c(12,13,14,15),1,0)
highsch = ifelse(person_direct$AGEP>=18 & person_direct$SCHL == 16,1,0)
somecol = ifelse(person_direct$AGEP>=18 & person_direct$SCH==18,1,0)
assobach = ifelse(person_direct$AGEP>=18 & person_direct$SCHL %in% c(20,21),1,0)
mastphd = ifelse(person_direct$AGEP>=18 & person_direct$SCHL %in% c(22,23,24),1,0)

lowage = ifelse(person_direct$AGEP <=5,1,0)
nosch = ifelse(person_direct$AGEP<18 & person_direct$SCHL %in% c(1, NA) ,1,0)
cnotstu = ifelse(person_direct$AGEP<18 & person_direct$SCH %in% c(1,NA),1,0)


hhid_temp = house_direct$hhid
temp = as.data.frame(cbind(hhid_temp,naemp))
hhid = as.data.frame(hhid)
temp1 = left_join(hhid,temp,by=c('hhid'='hhid_temp'))

loage1 = ifelse(lowage == 1 & temp1$naemp == 1, 1,0)
loage2 = ifelse(lowage == 1 & temp1$naemp == 2, 1,0)


income = ifelse(is.na(person_direct$PERNP),0, person_direct$PERNP/1000)

hhid_temp = house_direct$hhid
temp = as.data.frame(cbind(hhid_temp,hhincome))
hhid = as.data.frame(hhid)
temp1 = left_join(hhid,temp,by=c('hhid'='hhid_temp'))
incomef = ifelse(temp1$hhincome == 0, 0, income/temp1$hhincome)

Wdurlow = ifelse(person_direct$WKHP<20,1,0)
wdurmed = ifelse(person_direct$WKHP<40 & person_direct$WKHP>=20,1,0)
wdurhigh = ifelse(person_direct$WKHP>=40,1,0)
highglex = ifelse(person_direct$WKHP < 20, 1,0)
medflex = ifelse(person_direct$WKHP<40 & person_direct$WKHP>=20,1,0)
lowflex = ifelse(person_direct$WKHP>=40,1,0)

emptype1 = ifelse(aemp ==1 & loage1==1,0,0)
emptype2 = ifelse(aemp ==1 & loage1==1,0,0)
emptype3 = ifelse(aemp ==1 & loage1==1,0,0)
emptype4 = ifelse(aemp ==1 & loage1==1,0,0)
emptype5 = ifelse(aemp ==1 & loage1==1,1,0)
emptype6 = ifelse(aemp ==1 & loage1==1,1,0)
  

mother = ifelse(parent == 1 & female == 1, 1,0)
father = ifelse(parent == 1 & male == 1 ,1, 0)

aunemp = ifelse(person_direct$AGEP>=18 & aemp==0,1,0)

adchild = ifelse(person_direct$AGEP>=18 & person_direct$RELP == 2,1,0) 
adult = ifelse(person_direct$AGEP>=18,1,0)
wkhrs = person_direct$WKHP

placeholder = ifelse(loage1==1,0,0)

```

### combine all the person attr to table
```{r}
setwd("G:/CE-CERT/CEMDAP/test1/")
persons_generated =  as.data.frame(cbind(hhid,perid,aemp,stu,Lic,worktsz,schooltsz,female,age,parent,cauc,afamer,asian,male,
                                         pvehavbl,highglex,presch,kgtog4,g5tog8,g9orhigh,lowage,loage1,loage2,somecol,
                                         assobach,mastphd,income,incomef,Wdurlow,wdurmed,emptype1,emptype2,emptype3,emptype4,
                                         emptype5,mother,father,nosch,cnotstu,aunemp,adchild,adult,hisp,othrace,g8orlow,
                                         g9tog12,highsch,wkhrs,wdurhigh,medflex,lowflex,emptype6,placeholder,placeholder,
                                         placeholder,placeholder,placeholder,placeholder,placeholder))
persons_generated[is.na(persons_generated)] = 0
persons_generated=lapply(persons_generated,as.numeric)
dim(persons_generated)
write.table(persons_generated,"persons.dat",sep = "\t",col.names = F,row.names = F)
```



